---
title: "class08"
author: 'Lizzie (PID: 59010743)'
date: "10/22/2021"
output:
  pdf_document: default
  html_document: default
---

# Clustering methods

Kmeans clustering in R is done with the 'kmeans()' function.
Make up a data set to test.

```{r}
tmp <- c(rnorm(30,3), rnorm(30,-3))
data <- cbind(x=tmp, y=rev(tmp))
plot(data)
```

Run 'kmeans' set k to 2 nstart 20. You have to tell kmeans how many clusters you want.

```{r}
km <- kmeans(data, centers = 2, nstart = 20)
km
```
Q1. How many points are in each cluster?
```{r}
km$size
```
Q2. What component of your result object details cluster assignment/membership?
```{r}
km$cluster
```

Q3. What component of your result object details cluster center
```{r}
km$centers
```

Q4. Plot x colored by the kmeans cluster assignment and add cluster centers as blue points
```{r}
plot(data, col=km$cluster)
points(km$centers, col="blue", pch=15, cex=2)
```

#Hierarchical clustering using the hclust() function and the same data as before.You can't just pass the data to hclust. 
```{r}
hc <- hclust(dist(data))
hc

```

hclust has a plot method
```{r}
plot(hc)
```

To find out membership vector we need to "cut" the tree and for this you use the 'cutree()' function and tell it the height at which to cut it.

```{r}
cutree(hc, h=7)
```
We can also use the cutree() and state the number of k clusters we want
```{r}
grps <- cutree(hc, k=2)
```

```{r}
plot(data, col=grps)
```

Principal component analysis (PCA)


## PCA of UK food data
```{r}
url <- "https://tinyurl.com/UK-foods"
x <- read.csv(url)
dim(x)
x
#fix the data set
rownames(x) <- x[,1]
x <- x[,-1]
head(x)

#this isn't the best way. Deal upon import
url <- "https://tinyurl.com/UK-foods"
x <- read.csv(url, row.names=1)
x
```
```{r}
mycols <- rainbow(nrow(x))
barplot(as.matrix(x), col=rainbow(17), beside=TRUE)
```
```{r}
pairs(x, col=mycols, pch=16)
```
Here we will use the base R function for PCA which is called 'prcomp()'. You need to first transpose the data.
```{r}
pca <- prcomp(t(x))
summary(pca)
```
```{r}
plot(pca)
```

We want score plot (aka PCA plot). Basically a plot of PC1 vs. PC2
```{r}
attributes(pca)
```
We want the pca$x component for this plot
```{r}
plot(pca$x[,1:2])
text(pca$x[,1:2], labels = colnames(x))
```
We can examine PCA loadings which tell us how much the original variables contribute to each new PC.

```{r}
par(mar=c(10, 3, 0.35, 0))
barplot(pca$rotation[,1], las=2)
```

RNA Seq analysis
```{r}
url2 <- "https://tinyurl.com/expression-CSV"
rna.data <- read.csv(url2, row.names=1)
head(rna.data)
#how many genes
nrow(rna.data)
#how many experiments
ncol(rna.data)
colnames(rna.data)
```
```{r}
pca.rna <- prcomp(t(rna.data), scale=TRUE)
plot(pca$x[,1], pca$x[,2], xlab="PC1", ylab="PC2")
summary(pca.rna)
plot(pca.rna)
```
```{r}
plot(pca.rna$x[,1:2])
text(pca.rna$x[,1:2], labels = colnames(rna.data))
```

