---
title: "Class_17"
author: 'Lizzie (PID: 59010743)'
date: "11/23/2021"
output:
  pdf_document: default
  html_document: default
---

```{r}
vax <- read.csv("covid.csv")
head(vax)
test <- sort(vax$as_of_date, decreasing=TRUE)

```
Q1. What column details the total number of people fully vaccinated?
persons_fully_vaccinated
Q2. What column details the Zip code tabulation area? 
zip_code_tabulation_area
Q3. What is the earliest date in this dataset? 
2021-01-05
Q4. What is the latest date in this dataset? 
2021-11-16

```{r}
skimr::skim(vax)

```
Q5. How many numeric columns are in this dataset? 
9

Q6. Note that there are “missing values” in the dataset. How many NA values there in the persons_fully_vaccinated column? 8256
```{r}
sum(is.na(vax$persons_fully_vaccinated))
```

Q7. What percent of persons_fully_vaccinated values are missing (to 2 significant figures)? 
10.17%
```{r}
sum(is.na(vax$persons_fully_vaccinated))/nrow(vax) *100
```
```{r}
library(lubridate)
```

```{r}
today()
```
#Specify date format (ymd), then you can use the date column to, for example, figure out how many days have passed since the first data input.
```{r}
vax$as_of_date <- ymd(vax$as_of_date)
today() - vax$as_of_date[1]
vax$as_of_date[nrow(vax)] - vax$as_of_date[1]
```
Q9. How many days have passed since the last update of the dataset? 
7
```{r}
today()- vax$as_of_date[nrow(vax)]
```
Q10. How many unique dates are in the dataset (i.e. how many different dates are detailed)?
46
```{r}
length(unique(vax$as_of_date))
```
```{r}
library(zipcodeR)
```

```{r}
geocode_zip('92037')
zip_distance('92037','92109')
reverse_zipcode(c('92037', "92109") )
```


```{r}
# Pull data for all ZIP codes in the dataset
zipdata <- reverse_zipcode( vax$zip_code_tabulation_area )
```

```{r}
sd1 <- vax[ vax$county == "San Diego" , ]
head(sd1)
nrow(sd1)

library(dplyr)
sd <- filter(vax, county == "San Diego")
nrow(sd)
```

```{r}
sd.10 <- filter(vax, county == "San Diego" &
                age5_plus_population > 10000)
```

Q11. How many distinct zip codes are listed for San Diego County? 
107
```{r}
length(unique(sd$zip_code_tabulation_area))
```
Q12. What San Diego County Zip code area has the largest 12 + Population in this dataset?
92154
```{r}
which.max(sd$age12_plus_population)
sd$zip_code_tabulation_area[23]
```

Q13. What is the overall average “Percent of Population Fully Vaccinated” value for all San Diego “County” as of “2021-11-09
```{r}
sd_date <- filter(vax, county == "San Diego" & as_of_date == "2021-11-09")
head(sd_date)
summary(sd_date$percent_of_population_fully_vaccinated)
x <- sum(is.na(sd_date$percent_of_population_fully_vaccinated))
sum(na.omit(sd_date$percent_of_population_fully_vaccinated))/ (nrow(sd_date) - x) *100
```

Q14. Using either ggplot or base R graphics make a summary figure that shows the distribution of Percent of Population Fully Vaccinated values as of “2021-11-09”?
```{r}
hist(sd_date$percent_of_population_fully_vaccinated)
```
```{r}
ucsd <- filter(sd, zip_code_tabulation_area=="92037")
ucsd[1,]$age5_plus_population
```
```{r}
library(ggplot2)
```

Q15. Using ggplot make a graph of the vaccination rate time course for the 92037 ZIP code area:
```{r}
ucsd$as_of_date
ggplot(ucsd) +
  aes(as_of_date, 
  percent_of_population_fully_vaccinated) +
  geom_point() +
  geom_line(group=1) +
  ylim(c(0,1)) +
  labs(x="Date", y="Percent Vaccinated")
```

```{r}
# Subset to all CA areas with a population as large as 92037
vax.36 <- filter(vax, age5_plus_population > 36144 &
                as_of_date == "2021-11-16")
#head(vax.36)

```

Q16. Calculate the mean “Percent of Population Fully Vaccinated” for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2021-11-16”. Add this as a straight horizontal line to your plot from above with the geom_hline() function?
```{r}
mean(vax.36$percent_of_population_fully_vaccinated)
```

```{r}
ucsd$as_of_date
ggplot(ucsd) +
  aes(as_of_date, 
  percent_of_population_fully_vaccinated) +
  geom_point() +
  geom_line(group=1) +
  ylim(c(0,1)) +
  labs(x="Date", y="Percent Vaccinated") +
  geom_hline(yintercept = 0.6629812)
```
Q17. What is the 6 number summary (Min, 1st Qu., Median, Mean, 3rd Qu., and Max) of the “Percent of Population Fully Vaccinated” values for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2021-11-16”?
```{r}
dat<- summary(vax.36$percent_of_population_fully_vaccinated)
```
Q18. Using ggplot generate a histogram of this data.
```{r}
vax.36$percent_of_population_fully_vaccinated
ggplot(vax.36) +
  aes(vax.36$percent_of_population_fully_vaccinated)+
  geom_histogram()
```

Q19. Is the 92109 and 92040 ZIP code areas above or below the average value you calculated for all these above?

92109 = above, 92040 = below
```{r}
vax %>% filter(as_of_date == "2021-11-16") %>%  
  filter(zip_code_tabulation_area=="92040") %>%
  select(percent_of_population_fully_vaccinated)

vax %>% filter(as_of_date == "2021-11-16") %>%  
  filter(zip_code_tabulation_area=="92109") %>%
  select(percent_of_population_fully_vaccinated)
```
Q20. Finally make a time course plot of vaccination progress for all areas in the full dataset with a age5_plus_population > 36144.
```{r}
vax.36.all <- filter(vax, age5_plus_population > 36144)

ggplot(vax.36.all) +
  aes(as_of_date,
      percent_of_population_fully_vaccinated, 
      group=zip_code_tabulation_area) +
  geom_line(alpha=0.2, color="blue") +
  ylim(c(0,1)) +
  labs(x="Date", y="Percent Vaccinated",
       title="Vaccination Rate Across California",
       subtitle="Only areas with a population about 36K are shown") +
  geom_hline(yintercept = 0.6629812, linetype = "dashed")
```

