---
title: "Class_15"
author: 'Lizzie (PID: 59010743)'
date: "11/17/2021"
output:
  pdf_document: default
  html_document: default
---

```{r}
library(BiocManager)
library(DESeq2)
```

#Examining data published from Himes et al. 2014.

##Load the countData and colData
1. Count data is the count matrix (number of reads coming from each gene for each sample)
2. colData describes metadata about the columns of countData

```{r}
counts <- read.csv("airway_scaledcounts.csv", row.names=1)
metadata <- read.csv("airway_metadata.csv")
```

```{r}
head(counts)
head(metadata)
```

#Check that the first column of colData matches the column names of countData
```{r}
metadata$id == colnames(counts)
all(metadata$id == colnames(counts))
```

#Compare control to treated. First we need to access all the control columns in our counts data.
```{r}
treated <- metadata[metadata[,"dex"]=="treated",]
treated.counts <- counts[ ,treated$id]
treated.mean <- rowSums( treated.counts )/4 

control <- metadata[metadata[,"dex"]=="control",]
control.counts <- counts[ ,control$id]
control.mean <- rowSums( control.counts )/4 
head(control.mean)
```

```{r}
meancounts <- data.frame(control.mean, treated.mean)
nrow(counts)
```

```{r}
plot(meancounts[,1],meancounts[,2], xlab="Control", ylab="Treated", log="xy")
```

Transform the data to a log scale
```{r}
log2(80/20)
```

```{r}
meancounts$log2fc <- log2(meancounts[,"treated.mean"]/meancounts[,"control.mean"])
head(meancounts)
```
#Testing the which function. It is not useful in default mode on our type of multi-column input. Need to use arr.ind = TRUE
```{r}
which(c(T,F,T))
```
#This removes the zero values
```{r}
head(meancounts[,1:2])
zero.vals <- which(meancounts[,1:2]==0, arr.ind=TRUE)
to.rm <- unique(zero.vals[,1])
mycounts <- meancounts[-to.rm,]
head(mycounts)
nrow(mycounts)
```

```{r}
up.ind <- mycounts$log2fc > 2
sum(up.ind)/nrow(mycounts) * 100
down.ind <- mycounts$log2fc < (-2)
sum(down.ind)/nrow(mycounts) * 100
sum(up.ind, down.ind)/nrow(mycounts) * 100
```
#DESeq2 Analysis
#We first need to setup the DESeq input object
```{r}
dds <- DESeqDataSetFromMatrix(countData=counts, 
                              colData=metadata, 
                              design=~dex)
dds
dds <- DESeq(dds)
```

```{r}
res <- results(dds)
res
```

Volcano Plot
```{r}
plot( res$log2FoldChange,  -log(res$padj), 
      xlab="Log2(FoldChange)",
      ylab="-Log(P-value)")
abline(v=c(-2,2), col="pink", lty=2)
abline(h=-log(0.05), col="gray",lty=2)
```

#Add gene names to our dataset. Use the bioconductor packages. The first one does the work and the other contains the data.
```{r}
library("AnnotationDbi")
#BiocManager::install("org.Hs.eg.db")
library(org.Hs.eg.db)
```

#The main function we will use is mapIds()
```{r}
res$symbol <- mapIds(org.Hs.eg.db,
                     keys=row.names(res), # Our genenames
                     keytype="ENSEMBL",        # The format of our genenames
                     column="SYMBOL",          # The new format we want to add
                     multiVals="first")
write.csv(res, file = "allmyresults.csv")
```

