---
title: "Class06"
author: 'Lizzie (PID: 59010743)'
date: "10/15/2021"
output: pdf_document
---

## Rmarkdown Tutorial

**bold text**
*italic*

---

Code:

```{r}
plot(1:10)
```
## Function to Calculate Grade

```{r}
# input student grades
student1 <- c(100, 100, 100, 100, 100, 100, 100, 90)  
student2 <- c(100, NA, 90, 90, 90, 90, 97, 80)  
student3 <- c(90, NA, NA, NA, NA, NA, NA, NA)
```

Find the minimum score using **min()** and where its position in the vector using **which.min()**

```{r}
which.min(student1)
```

```{r}
student1[-which.min(student1)]
```

Find the mean of the vector with the lowest score removed.
```{r}
mean(student1[-which.min(student1)])
```

The above works if there are no NA's in the vector. The is.na() function returns a logical vector where the TRUE elements indicate and NA. A "!" means not.

```{r}
is.na(student2)
```

Replace the NAs with zero nad insert previous code
```{r}
student2.prime <- student2
student2.prime[is.na(student2.prime)]=0
student2.prime
mean(student2.prime[-which.min(student2.prime)])

```

Check to see if it works with student 3
```{r}
student3.prime <- student3
student3.prime[is.na(student3.prime)]=0
student3.prime
mean(student3.prime[-which.min(student3.prime)])
```

What if one of the entries is as a string? Use **as.numeric()**
```{r}
student4 <- c(100, NA, 90, "90", 90, 90, 97, 80)
x <- as.numeric(student4)
x[is.na(x)]=0
mean(x[-which.min(x)])
```
Write the function
```{r}
grade <- function(x){
  x <- as.numeric(x)
  x[is.na(x)]=0
  mean(x[-which.min(x)])
}
```

Test Function
```{r}
grade(student1)
```
## Now grade a whole class

First we read the gradebook from the class
```{r}
gradebook <- "https://tinyurl.com/gradeinput"
scores <- read.csv(gradebook, row.names=1)
scores

```


Use the function **apply()** to grade all students with our **grade()** function
```{r}
ans <- apply(scores, 1, grade)
```

Q2: Who is the top scoring student?
```{r}
which.max(ans)
```

Q3: Which homework was the toughest? 
```{r}
hw <- apply(scores, 2, mean)
hw
```

Remove the NAs
```{r}
mask <- scores
mask[is.na(mask)]=0
mask
```

Q3: Which homework was the toughest? 
```{r}
worst <- apply(mask, 2, mean)
worst
```

Q4: What score was the most predictive of overall score? Here we will use the **cor()** function
```{r}
cor(mask$hw1, ans)
cor(mask$hw2, ans)
cor(mask$hw3, ans)
cor(mask$hw4, ans)
cor(mask$hw5, ans)
```
Call for **cor()** function for every homework using the **apply()** function. Hwk 5 is the most predictive.
```{r}
apply(mask,2,cor,ans)
```

Make a boxplot
```{r}
boxplot(scores)
```
```{r}
grade <-function(x, drop.lowest = TRUE){
x <- as.numeric(x)
x[is.na(x)]=0
if (drop.lowest){
  mean(x[-which.min(x)])
} else{
  mean(x)
}
```}

